---
title: "Glucose"
output: html_notebook
---

#Setup

```{r}
require(data.table)
require(ggplot2)
require(table1)
require(readr)
library(olsrr)
library(faraway)
library(caret)
library(dplyr)
library(Hmisc)
```

# data

```{r}
dt1 <- read.csv("glucose.csv")
head(dt1)
summary(dt1)
```

## Clean data

```{r}
dt1$Ins_Admin <- dt1$Insulin > 0
dt1$Outcome <- NULL
dt1 <- dt1[dt1$Glucose > 0 &
             dt1$BMI > 0 & dt1$BloodPressure > 0, ]

summary(dt1)
hist(dt1$Glucose, 100)
hist(log(dt1$Glucose), 100)
hist(dt1$BMI, 100)
hist(log(dt1$BMI), 100)

```

# Plot histogram

```{r}
hist.data.frame(dt1)
```

# Table 1

```{r}
table1(~. | Ins_Admin,
       data = dt1, topclass="Rtable1-grid")

```

# Predictors of glucose

## Insulin

```{r}
plot(dt1$Glucose ~ dt1$Insulin)
plot(dt1$Glucose ~ log(dt1$Insulin))
# plot(log(dt1$Glucose) ~ log(dt1$Insulin))

dt1$sqrtInsulin <- sqrt(dt1$Insulin)

summary(lm(Glucose ~ sqrtInsulin,
           data = dt1))
```
## Blood pressure
```{r}
plot(dt1$Glucose ~ dt1$BloodPressure)
plot(dt1$Glucose ~ log(dt1$BloodPressure))

m1 <- loess(Glucose ~ BloodPressure,
            data = dt1)
dt1$bp_pred <- predict(m1)

ggplot(dt1,
       aes(x = BloodPressure,
           y =  Glucose))+
  geom_point() +
  geom_line(aes(x = BloodPressure,
                y = bp_pred)) +
  theme_bw()

dt1$bp_under_70 <- dt1$BloodPressure < 70

summary(lm(Glucose ~ BloodPressure,
           data = dt1))
summary(lm(Glucose ~ bp_under_70,
           data = dt1))
```
## Age
```{r}
# plot(dt1$Glucose ~ dt1$Age)
# plot(dt1$Glucose ~ log(dt1$Age))
# plot(log(dt1$Glucose) ~ dt1$Age)

m2 <- loess(Glucose ~ Age,
      data = dt1)

dt1$gl_pred <- predict(m2)

ggplot(dt1,
       aes(x = Age,
           y = Glucose)) +
  geom_point() +
  geom_line(aes(x = Age,
                y = gl_pred)) +
  theme_bw()

dt1$age_under_30 <- dt1$Age < 30

summary(lm(Glucose ~ Age,
   data = dt1))
summary(lm(Glucose ~ age_under_30,
   data = dt1))
```

## DiabetsPedigreeFunction

```{r}
plot(dt1$Glucose ~ dt1$DiabetesPedigreeFunction)
plot(dt1$Glucose ~ log(dt1$DiabetesPedigreeFunction))

m3 <- loess(Glucose ~ DiabetesPedigreeFunction,
            data = dt1)
dt1$dpf_pred <- predict(m3)

ggplot(dt1,
       aes(x = DiabetesPedigreeFunction,
           y =  Glucose))+
  geom_point() +
  geom_line(aes(x = DiabetesPedigreeFunction,
                y = dpf_pred)) +
  theme_bw()


summary(lm(Glucose ~ DiabetesPedigreeFunction,
           data = dt1))
summary(lm(Glucose ~ log(DiabetesPedigreeFunction),
           data = dt1))
```

## BMI

```{r}
plot(dt1$Glucose ~ dt1$BMI)
plot(dt1$Glucose ~ log(dt1$BMI))
plot(log(dt1$Glucose) ~ log(dt1$BMI))

m4 <- loess(log(Glucose)~ log(BMI), data = dt1)
ggplot(dt1,
       aes(y = log(Glucose),
           x = log(BMI))) + 
  geom_point() + 
  geom_line(aes(x = log(BMI),
                y = predict(m4))) + 
  theme_bw()
  
summary(m4)
```

## Pregnancies

```{r}
plot(dt1$Glucose ~ dt1$Pregnancies)
plot(dt1$Glucose ~ log(dt1$Pregnancies))
plot(log(dt1$Glucose) ~ log(dt1$Pregnancies))

m5 <- loess(Glucose~Pregnancies, data = dt1)
ggplot(dt1,
       aes(y = Glucose,
           x = Pregnancies)) + 
  geom_point() + 
  geom_line(aes(x = Pregnancies,
                y = predict(m5))) + 
  theme_bw()

summary(m5)
```
## SkinThickness
```{r}
plot(dt1$Glucose ~ dt1$SkinThickness)
plot(dt1$Glucose ~ log(dt1$SkinThickness))
plot(log(dt1$Glucose) ~ log(dt1$SkinThickness))

m6 <- lm(Glucose ~ SkinThickness,
           data = dt1)
summary(m6)
```

### Skin tikness of patients who are not diabetic wasn't mesured, so we will dropp this variable to avoid any incorrect results

```{r}
dt1$SkinThickness <- NULL
```


```{r}
pairs(~ log(Glucose) + Pregnancies + BloodPressure + Insulin + BMI + 
      DiabetesPedigreeFunction + Age,
      data = dt1, panel = panel.smooth,
      main = "Glucose data factors",
      pch = 21,cex = 0.8,
      bg = c("red", "green3", "blue")[unclass(iris$Species)])
```

# Linear Regression Model

```{r}
m0 <- lm(log(Glucose) ~ Pregnancies + BloodPressure +
           Insulin + BMI + DiabetesPedigreeFunction + Age , data = dt1)
summary(m0)
```

# To select variables that were significantly associated with the outcome (plasma glucose concentration) we will test stepwise forward,backward, by direction and best-subset procedures

## All best subset selection

```{r}
k1 <- ols_step_best_subset(m0)
k1
```

### Model 5 has higher Adj. R sq BloodPressure Insulin BMI DiabetesPedigreeFunction Age

## Stepwise forward selection

```{r}
k2 <- ols_step_forward_p(m0, penter = 0.05)
k2
```

## Stepwise backward selection

```{r}
k3 <- ols_step_backward_p(m0)
k3$model
```

## By direction selection

```{r}
k4 <- ols_step_both_p(m0)
k4$predictors
```

# Will perform a 5-fold Cross-Validation comparing model from foorward with model by direction
## Step forward selected model
```{r}
lmod1 <- lm(Glucose ~ Insulin + Age +
              BloodPressure +BMI +  DiabetesPedigreeFunction, dt1)

set.seed(13245)
train.control_1 <- trainControl(method = "cv", number = 5)
model_1 <- train(Glucose ~ Insulin + Age +
                   BloodPressure +BMI + DiabetesPedigreeFunction,
                 data = dt1 ,method = "lm", trControl = train.control_1)
print(model_1$results)
summary(lmod1)
```

## By direction selected model
```{r}
lmod2 <- lm(Glucose ~ Insulin + Age + BloodPressure + BMI, data = dt1)

set.seed(13245)
train.control_2 <- trainControl(method = "cv", number = 5)
model_2 <- train(Glucose ~ Insulin + Age
                 + BloodPressure + BMI, data = dt1,
                 method = "lm", trControl = train.control_2)
print(model_2$results)
summary(lmod2)

```
```{r}
anova(model_1, model_2)
```




#lmod_1      RMSE      Rsquared   MAE
#           24.45051  0.3917216  18.65624

#model1      RSE      MRS      Adj_RS   MAE 
#           24.49     0.3995   0.3917 



#lmod_2      RMSE      Rsquared   MAE    
#           24.45103  0.3912515  18.6361

#model2    RSE      MRS      Adj_RS   MAE 
#           24.54     0.394   0.3893
```