---
title: "Constructing a Predictive Model for Plasma Glucose Concentration in Female Patients"
output:
  html_notebook:
    highlight: tango
    number_sections: yes
    theme: united
    toc: TRUE
    toc_float: TRUE
    df_print: paged
    code_folding: hide
---

# Author: Vahe Nersisyan    
Date: `r date()`


# Load libraries
```{r}
require(data.table)
require(ggplot2)
require(table1)
require(readr)
library(olsrr)
library(caret)
library(dplyr)
library(psych)
library(Hmisc)
```

# Data
```{r}
dt1 <- read.csv("glucose.csv")
head(dt1)
summary(dt1)
```
## Clean data
```{r}
dt1$Outcome <- NULL
dt1 <- dt1[dt1$Glucose > 0 &
             dt1$BMI > 0 & dt1$BloodPressure > 0, ]
dt1$SkinThickness <-  NULL
summary(dt1)
```
## Plot data histogram
```{r}
hist.data.frame(dt1)
```

## Scaterplot
```{r}
pairs(Glucose ~ Pregnancies + BloodPressure + Insulin + BMI + 
      DiabetesPedigreeFunction + Age,
      data = dt1, panel = panel.smooth,
      main = "Simple scaterplot ",
      pch = 21,cex = 0.8,
      bg = c("red", "green3", "blue")[unclass(iris$Species)])
```

# Predictors of glucose
## Insulin
```{r}
dt1$Ins_Admin <- dt1$Insulin > 0
plot(dt1$Glucose ~ dt1$Insulin)
plot(dt1$Glucose ~ log(dt1$Insulin))
plot(log(dt1$Glucose) ~ log(dt1$Insulin))
plot(log(dt1$Glucose) ~ dt1$Insulin)

summary(lm(Glucose ~ Insulin,
           data = dt1))

dt1$Ins_Admin <- factor(dt1$Ins_Admin, levels = c(TRUE, FALSE), labels = c("Diabetic", "Non-Diabetic"))
cor(dt1$Glucose, dt1$Insulin)

```
## Blood pressure
```{r}
plot(dt1$Glucose ~ dt1$BloodPressure)
plot(dt1$Glucose ~ log(dt1$BloodPressure))

m1 <- loess(Glucose ~ BloodPressure,
            data = dt1)
dt1$bp_pred <- predict(m1)

ggplot(dt1,
       aes(x = BloodPressure,
           y =  Glucose))+
  geom_point() +
  geom_line(aes(x = BloodPressure,
                y = bp_pred)) +
  theme_bw()

dt1$bp_under_70 <- dt1$BloodPressure < 70

summary(lm(Glucose ~ BloodPressure,
           data = dt1))
summary(lm(Glucose ~ bp_under_70,
           data = dt1))
```

## Age
```{r}
plot(dt1$Glucose ~ dt1$Age)
plot(dt1$Glucose ~ log(dt1$Age))
plot(log(dt1$Glucose) ~ dt1$Age)

m2 <- loess(Glucose ~ Age,
      data = dt1)

dt1$age_pred <- predict(m2)

ggplot(dt1,
       aes(x = Age,
           y = Glucose)) +
  geom_point() +
  geom_line(aes(x = Age,
                y = age_pred)) +
  theme_bw()


summary(lm(Glucose ~ Age,
   data = dt1))
dt1$age_under_30 <- dt1$Age < 30
summary(lm(Glucose ~ age_under_30,
   data = dt1))

dt1$age_under_30 <- factor(dt1$age_under_30, levels = c(TRUE, FALSE), labels = c("Under 30", "Over 30"))


```

## DiabetsPedigreeFunction
```{r}
plot(dt1$Glucose ~ dt1$DiabetesPedigreeFunction)
plot(dt1$Glucose ~ log(dt1$DiabetesPedigreeFunction))

m3 <- loess(Glucose ~ DiabetesPedigreeFunction,
            data = dt1)
dt1$dpf_pred <- predict(m3)

ggplot(dt1,
       aes(x = DiabetesPedigreeFunction,
           y =  Glucose))+
  geom_point() +
  geom_line(aes(x = DiabetesPedigreeFunction,
                y = dpf_pred)) +
  theme_bw()


summary(lm(Glucose ~ DiabetesPedigreeFunction,
           data = dt1))
summary(lm(Glucose ~ log(DiabetesPedigreeFunction),
           data = dt1))
```

## BMI
```{r}
plot(dt1$Glucose ~ dt1$BMI)
plot(dt1$Glucose ~ log(dt1$BMI))
plot(log(dt1$Glucose) ~ log(dt1$BMI))

m4 <- loess(log(Glucose)~ log(BMI), data = dt1)
ggplot(dt1,
       aes(y = log(Glucose),
           x = log(BMI))) + 
  geom_point() + 
  geom_line(aes(x = log(BMI),
                y = predict(m4))) + 
  theme_bw()
  
summary(m4)
```

## Pregnancies
```{r}
plot(dt1$Glucose ~ dt1$Pregnancies)
plot(dt1$Glucose ~ log(dt1$Pregnancies))
plot(log(dt1$Glucose) ~ log(dt1$Pregnancies))

m5 <- loess(Glucose~Pregnancies, data = dt1)
ggplot(dt1,
       aes(y = Glucose,
           x = Pregnancies)) + 
  geom_point() + 
  geom_line(aes(x = Pregnancies,
                y = predict(m5))) + 
  theme_bw()

summary(m5)
```

# Table 1
```{r}
# Age Table
table1(~ Glucose + Pregnancies + BloodPressure +
         Insulin + BMI + DiabetesPedigreeFunction| age_under_30,
       data = dt1, topclass="Rtable1-grid")

# Insulin Table
table1(~ Glucose + Pregnancies + BloodPressure + BMI +  
         DiabetesPedigreeFunction + Age | Ins_Admin, 
       data = dt1, topclass="Rtable1-grid")

```

# Full Linear Regression Model
```{r}
dt1$Glucose <- log(dt1$Glucose)
lmod1 <- lm(Glucose ~ Pregnancies + BloodPressure +
           Insulin + BMI + DiabetesPedigreeFunction +
             Age , data = dt1)


summary(lmod1)
plot(lmod1)

```

# Model selection

## All best subset selection
```{r}
step_best <- ols_step_best_subset(lmod1)
step_best
plot(step_best)
```

###Model 5 has higher Adj. R sq BloodPressure Insulin BMI DiabetesPedigreeFunction Age

## Stepwise forward selection
```{r}
step_frw <- ols_step_forward_p(lmod1, penter = 0.05)
step_frw
plot(step_frw)
```
## Stepwise backward selection
```{r}
step_bkw <- ols_step_backward_p(lmod1, prem = 0.05)
step_bkw
plot(step_bkw)
```

## By direction selection
```{r}
step_bydir <- ols_step_both_p(lmod1, pent = 0.05, prem = 0.05)
step_bydir
plot(step_bydir)
```

# 5 Fold Cross Validation
##Performing 5-fold Cross-Validation comparing best subset selected model with step forward selecting model
### Best subset selected model
```{r}
model_best <- lm(Glucose ~ BloodPressure + Insulin +
                   BMI + DiabetesPedigreeFunction + Age,
                 data = dt1)


set.seed(13245)
train.control_1 <- trainControl(method = "cv", number = 5)
model_best_cross <- train(Glucose ~ BloodPressure + Insulin +
                            BMI + DiabetesPedigreeFunction + Age,
                          data = dt1, method = "lm",
                          trControl = train.control_1)

print(model_best)
summary(model_best_cross)
```
### Step forward selected model
```{r}
model_fwd <- lm(Glucose ~ Insulin + Age +
                  BloodPressure + BMI, data = dt1)


set.seed(13245)
train.control_2 <- trainControl(method = "cv", number = 5)
model_fwd_cross <- train(Glucose ~ Insulin + Age
                         + BloodPressure + BMI, data = dt1,
                         method = "lm", trControl = train.control_2)

print(model_fwd)
summary(model_fwd_cross)

```
### Using model_best_cross and model_fwd_cross to see if there are any difference in metrics for 5-fold cross validation with and without repeats

```{r}
train.control2 <- trainControl(method = "repeatedcv",
                               number = 5,repeats=6)

model_fwd_cross_2 <- train(Glucose ~ Insulin + Age + 
                             BloodPressure + BMI, data = dt1, 
                           method = "lm", trControl = train.control2)

print(model_fwd_cross_2)


train.control1 <- trainControl(method = "repeatedcv",
                               number = 5,repeats=6)

model_best_cross_2 <- train(Glucose ~ BloodPressure + Insulin + BMI +
                              DiabetesPedigreeFunction + Age,
                            data = dt1, method = "lm",
                            trControl = train.control1)

print(model_best_cross_2)

```

# Testing model
```{r}
ols_test_correlation(model_best)
ols_test_normality(model_best)
ols_plot_hadi(model_best)
ols_plot_resid_box(model_best)
ols_plot_diagnostics(model_best)
```
###The value shows that our predictors are normally distributed

## Finding CI
```{r}
confint(lmod1)
```

# Test selected model
```{r}
ols_coll_diag(lmod1)
ols_correlations(lmod1)
ols_plot_diagnostics(lmod1)
```


